### 2019-07-05 manuscript/manuscript.tex
Update manuscript structure and organization
The new organization reflects the plan for the manuscript outline and the
structure more closely aligns this paper with others from the lab.

\documentclass[12pt]{article}

% graphicx package, useful for including eps and pdf graphics
\usepackage{graphicx}
\DeclareGraphicsExtensions{.png,.png,.jpg}

% basic packages
\usepackage{color}
\usepackage{parskip}
\usepackage{float}
\usepackage{microtype}
\usepackage{url}
\usepackage{hyperref}

% text layout
\usepackage{geometry}
\geometry{textwidth=17cm} % 15.25cm for single-space, 16.25cm for double-space
\geometry{textheight=22.5cm} % 22cm for single-space, 22.5cm for double-space

% helps to keep figures from being orphaned on a page by themselves
\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}

% bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,font=small]{caption}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{authblk}
\renewcommand\Authands{ \& }
\renewcommand\Authfont{\normalsize \bf}
\renewcommand\Affilfont{\small \normalfont}
\makeatletter
\renewcommand\AB@affilsepx{, \protect\Affilfont}
\makeatother

\title{Integrative forecasting of seasonal influenza A/H3N2 by genotype and phenotype}
%\title{Experimentally informed forecasts of seasonal influenza A/H3N2}
%\title{Long-term forecasts of seasonal influenza reveal historical contingency of fitness metrics}

\author[1,2]{John Huddleston}
\author[2]{Richard A.\ Neher}
\author[1]{Trevor Bedford}

\affil[1]{Vaccine and Infectious Disease Division, Fred Hutchinson Cancer Research Center, Seattle, WA, USA}
\affil[2]{Moleculary and Cell Biology, University of Washington, Seattle, WA, USA}
\affil[3]{Biozentrum, University of Basel, Basel, Switzerland}

\begin{document}

\begin{abstract}
\input{abstract.tex}
\end{abstract}

\maketitle

\input{main.tex}
\bibliographystyle{nih}
\bibliography{manuscript}

\clearpage

\input{supplement.tex}

\end{document}