### 2019-11-07 manuscript/supplement.tex
Simplify captions, update natural population tables and figures

\setcounter{figure}{0}
\setcounter{table}{0}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\Roman{table}}

\section*{Supplemental Material}

\begin{figure*}[h]
  \begin{center}
  \includegraphics[width=\textwidth]{figures/cross-validation-for-simulated-populations.png}
  \caption{
  Time-series cross-validation scheme for simulated populations.
  Models were trained in six-year sliding windows (grey lines) and validated on out-of-sample data from validation timepoints (filled circles).
  Validation results from 30 years of data were used to iteratively tune model hyperparameters.
  After fixing hyperparameters, model coefficients were fixed at the mean values across all training windows.
  Fixed coefficients were applied to 10 years of new out-of-sample test data (open circles) to estimate true forecast errors.
  }
  \label{sup_fig:cross_validation_for_simulated_populations}
  \end{center}
\end{figure*}

\begin{figure*}[h]
  \begin{center}
  \includegraphics[width=\textwidth]{figures/simulated-h3n2-ha-phylogeny.png}
  \caption{
  Phylogeny of A/H3N2-like HA sequences sampled between the 24th and 30th years of simulated evolution.
  The phylogenetic structure and rate of accumulated epitope and non-epitope mutations match patterns observed in phylogenies of natural sequences.
  Sample dates were annotated as the generation in the simulation divided by 200 and added to 2000, to acquire realistic date ranges that were compatible with our modeling machinery.
  }
  \label{sup_fig:simulated_h3n2_ha_phylogeny}
  \end{center}
\end{figure*}

\begin{figure*}[t]
  \begin{center}
  \includegraphics[width=\textwidth]{figures/unadjusted-composite-model-accuracy-and-coefficients-for-simulated-populations.png}
  \caption{Composite model a) accuracy and b) coefficients for simulated populations of A/H3N2-like viruses.}
  \label{sup_fig:unadjusted_composite_model_accuracy_and_coefficients_for_simulated_populations}
  \end{center}
\end{figure*}

\begin{figure*}[t]
  \begin{center}
  \includegraphics[width=\textwidth]{figures/validation-of-best-model-for-simulated-populations.png}
  \caption{
  Validation of best model for simulated populations of A/H3N2-like viruses.
  a) The correlation of estimated and observed clade growth rates shows the model's ability to capture clade-level dynamics without explicitly optimizing for clade frequency targets.
  b) The rank of the estimated best strain based on its distance to the future in the best model shows how often the model makes a good choice when forced to select a single representative strain for the future population.
  }
  \label{sup_fig:validation_of_best_model_for_simulated_populations}
  \end{center}
\end{figure*}

\begin{figure*}[h]
  \begin{center}
  \includegraphics[width=\textwidth]{figures/cross-validation-for-natural-populations.png}
  \caption{
  Time-series cross-validation scheme for natural populations.
  Models were trained in six-year sliding windows (grey lines) and validated on out-of-sample data from validation timepoints (filled circles).
  Validation results from 25 years of data were used to iteratively tune model hyperparameters.
  After fixing hyperparameters, model coefficients were fixed at the mean values across all training windows.
  Fixed coefficients were applied to four years of new out-of-sample test data (open circles) to estimate true forecast errors.
  }
  \label{sup_fig:cross_validation_for_natural_populations}
  \end{center}
\end{figure*}

\begin{table*}[ht]
  \begin{center}
    \input{tables/complete_natural_model_selection.tex}
    \caption{
      All model coefficients and validation performance for natural populations ordered from best to worst by distance to the future, as in Table~\ref{table_simulated_model_selection}.
      Results are based on training and validation across 23 training windows.
      Model results for additional variants of fitness metrics including those based on epitope mutations and DMS preferences are included for reference.
    }
    \label{sup_table:complete_natural_model_selection}
  \end{center}
\end{table*}

\begin{table*}[ht]
  \begin{center}
    \input{tables/mutations_by_trunk_status.tex}
    \caption{
    Number of epitope and non-epitope mutations per branch by trunk or side branch status for natural populations.
    Epitope sites were defined previously described \cite{Luksza:2014hj}.
    Annotation of trunk and side branch was performed as previously described \cite{Bedford:2015fj}.
    Mutations were calculated for the full validation tree for natural sequences samples between 1990 and 2015.
    }
    \label{sup_table:mutations_by_trunk_status}
  \end{center}
\end{table*}

\begin{table*}[ht]
  \begin{center}
    \input{tables/mutations_by_trunk_status_for_simulated_populations.tex}
    \caption{
    Number of epitope and non-epitope mutations per branch by trunk or side branch status for simulated populations.
    Epitope sites were defined previously described \cite{Luksza:2014hj}.
    Annotation of trunk and side branch was performed as previously described \cite{Bedford:2015fj}.
    Mutations were calculated for the full validation tree for simulated sequences samples between October of years 10 and 40.
    }
    \label{sup_table:mutations_by_trunk_status_for_simulated_populations}
  \end{center}
\end{table*}

\begin{figure*}[t]
  \begin{center}
  \includegraphics[width=\textwidth]{figures/unadjusted-DMS-model-accuracy-and-coefficients-for-natural-populations.png}
  \caption{
  Primary and alternate fitness metrics based on deep mutational scanning (DMS) preferences for seasonal influenza A/H3N2 strain A/Perth/16/2009.
  Metrics are increasingly less dependent on the sequence content of the DMS strain's genetic background from top to bottom.
  The most structurally naive metric of non-epitope mutations outperforms all of the DMS metrics, indicating that these metrics still overfit to the DMS strain's genetic background.
  }
  \label{sup_fig:unadjusted_DMS_model_accuracy_and_coefficients_for_natural_populations}
  \end{center}
\end{figure*}