### 2020-01-02 manuscript/abstract.tex
Edit abstract to 150 words or less
This is 149 words according to emacs word count and 144 according to wc.

Seasonal influenza virus A/H3N2 is a major cause of death globally.
Vaccination remains the most effective preventative.
Rapid mutation of hemagglutinin allows viruses to escape adaptive immunity.
This antigenic drift necessitates regular vaccine updates.
Effective vaccine strains need to match H3N2 populations circulating one year after strain selection.
Experts select strains based on experimental measurements of antigenic drift and predictions made by models from hemagglutinin sequences.
No modern models use phenotypic measures of viral fitness in their predictions.
We developed a novel influenza forecasting framework that integrates phenotypic measures of antigenic drift and functional constraint with previously published sequence-only fitness estimates.
Forecasts informed by phenotypic measures of antigenic drift consistently outperformed previous sequence-only estimates, while sequence-only estimates of functional constraint surpassed more comprehensive experimentally-informed estimates.
Importantly, the best models integrated estimates of both functional constraint and either antigenic drift phenotypes or recent population growth.