### 2020-06-06 manuscript/tables/composite_vs_individual_model_comparison.tex
Add bootstrap comparisons between composite and individual models

\begin{tabular}{llllrr}
\toprule
    sample &  error\_type &      individual\_model &                         composite\_model &  empirical\_difference &  p\_value \\
\midrule
 simulated &  validation &          true fitness &                  mutational load +  LBI &                  0.42 &   0.9678 \\
 simulated &  validation &       mutational load &                  mutational load +  LBI &                 -1.03 &   0.0000 \\
 simulated &  validation &                   LBI &                  mutational load +  LBI &                 -0.33 &   0.0096 \\
 simulated &        test &          true fitness &                  mutational load +  LBI &                 -0.28 &   0.1421 \\
 simulated &        test &       mutational load &                  mutational load +  LBI &                 -1.11 &   0.0000 \\
 simulated &        test &                   LBI &                  mutational load +  LBI &                 -0.41 &   0.0002 \\
   natural &  validation &       mutational load &                  mutational load +  LBI &                 -0.69 &   0.0046 \\
   natural &  validation &                   LBI &                  mutational load +  LBI &                 -0.23 &   0.0027 \\
   natural &  validation &       mutational load &  mutational load + HI antigenic novelty &                 -0.31 &   0.0410 \\
   natural &  validation &  HI antigenic novelty &  mutational load + HI antigenic novelty &                 -0.18 &   0.0559 \\
   natural &        test &       mutational load &                  mutational load +  LBI &                  1.17 &   0.9400 \\
   natural &        test &                   LBI &                  mutational load +  LBI &                 -0.70 &   0.0000 \\
   natural &        test &       mutational load &  mutational load + HI antigenic novelty &                 -0.56 &   0.0135 \\
   natural &        test &  HI antigenic novelty &  mutational load + HI antigenic novelty &                 -0.24 &   0.0948 \\
\bottomrule
\end{tabular}