### 2020-06-09 manuscript/tables/composite_vs_individual_model_comparison.tex
Report effect sizes for bootstrap analyses
Adds mean +/- standard deviation of bootstrap distributions to violin plots of
distributions to better communicate the effect sizes of model improvements over
the naive.

\begin{tabular}{llllrr}
\toprule
    sample &  error\_type &      individual\_model &                         composite\_model &  empirical\_difference &  p\_value \\
\midrule
 simulated &  validation &          true fitness &                  mutational load +  LBI &                  0.42 &   0.9670 \\
 simulated &  validation &       mutational load &                  mutational load +  LBI &                 -1.03 &   0.0000 \\
 simulated &  validation &                   LBI &                  mutational load +  LBI &                 -0.33 &   0.0098 \\
 simulated &        test &          true fitness &                  mutational load +  LBI &                 -0.28 &   0.1391 \\
 simulated &        test &       mutational load &                  mutational load +  LBI &                 -1.11 &   0.0000 \\
 simulated &        test &                   LBI &                  mutational load +  LBI &                 -0.41 &   0.0001 \\
   natural &  validation &       mutational load &                  mutational load +  LBI &                 -0.69 &   0.0028 \\
   natural &  validation &                   LBI &                  mutational load +  LBI &                 -0.23 &   0.0037 \\
   natural &  validation &       mutational load &  mutational load + HI antigenic novelty &                 -0.31 &   0.0396 \\
   natural &  validation &  HI antigenic novelty &  mutational load + HI antigenic novelty &                 -0.18 &   0.0534 \\
   natural &        test &       mutational load &                  mutational load +  LBI &                  1.17 &   0.9365 \\
   natural &        test &                   LBI &                  mutational load +  LBI &                 -0.70 &   0.0000 \\
   natural &        test &       mutational load &  mutational load + HI antigenic novelty &                 -0.56 &   0.0133 \\
   natural &        test &  HI antigenic novelty &  mutational load + HI antigenic novelty &                 -0.24 &   0.0965 \\
\bottomrule
\end{tabular}