### 2020-07-29 manuscript/tables/model_summary.tex
Replace reference to this study with "this study"

\begin{tabular*}{1.4\textwidth}{lllll}
\toprule
Model                                                        & Populations        & Data type             & Fitness category      & Originally implemented by                   \\
\midrule
true fitness                                                 & simulated          & simulated populations & positive control      & this study                      \\
naive                                                        & simulated, natural & HA sequences          & negative control      & this study                      \\
epitope antigenic novelty                                    & simulated, natural & HA sequences          & antigenic drift       & {\L}uksza and L\"assig \cite{Luksza:2014hj} \\
epitope ancestor                                             & simulated, natural & HA sequences          & antigenic drift       & {\L}uksza and L\"assig \cite{Luksza:2014hj} \\
HI antigenic novelty                                         & natural            & serological assays    & antigenic drift       & this study                      \\
mutational load                                              & simulated, natural & HA sequences          & functional constraint & {\L}uksza and L\"assig \cite{Luksza:2014hj} \\
\makecell{deep mutational scanning (DMS) \\ mutational effects} & natural            & DMS assays            & functional constraint & Lee et al. 2018 \cite{Lee2018}              \\
local branching index (LBI)                                  & simulated, natural & HA sequences          & clade growth          & Neher et al. 2014 \cite{Neher:2014eu}       \\
delta frequency                                              & simulated, natural & HA sequences          & clade growth          & this study                      \\
\bottomrule
\end{tabular*}